---
title: Compile equivalence table for suicide methods in ICD-9 and ICD-10 (intentional
  & undetermined intent)
author: "Jan Savinc"
date: "04/09/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


# Introduction

For compiling suicide statistics, we conventionally include external cause codes both for intentional self-injury and for undetermined intent. In order to group longitudinal results, we need to compile equivalent codes in ICD-9 and ICD-10, and equivalent methods of injury between intentional and undetermined intent codes.

The codes include:

* intentional self-harm (ICD-9 codes E950-959; ICD-10 codes X60-X84 plus
Y87.0 (sequelae of)
* events of undetermined intent (ICD-9 codes E980-989; ICD-10 codes Y10-
Y34 plus Y87.2 (sequelae of)

Of those, ICD-10 codes X60-X84 map one-to-one to codes Y10-Y34 in terms of cause or means of injury; ICD-9 codes E950-E959 map onto E980-E989.
The same is true for Y87.0, Y87.2, and E959 & E989 (sequelae of self-harm or undetermined intent)

The trickier task is to map ICD-9 codes onto ICD-10 codes - whereas ICD-10 codes are more spread-out, and in death records mostly (potentially always) appear as 3-character codes, ICD-9 codes use 4-th digit distinctions.

For example, self-poisonings are X60-X69 in ICD-10, whereas in ICD-9 self-poisonings are in the range E950-E952.

# Reusing previous groupings

The grouping provided by the [http://mchp-appserv.cpe.umanitoba.ca/viewConcept.php?conceptID=1183](Manitoba Centre for Health Policy):

* poisoning with undetermined intent: ICD-10 codes Y10-Y12, Y16, Y17
* self-inflicted poisoning: ICD-9 codes E950-E952, ICD-10 codes X60-X69
* self-inflicted injury by hanging, strangulation and suffocation: ICD-9 code E953, ICD-10 code X70
* self-inflicted injury by drowning: ICD-9 code E954, ICD-10 code X71
* self-inflicted injury by firearms and explosives: ICD-9 code E955, ICD-10 codes X72-X75
* self-inflicted injury by smoke, fire, flames, steam, hot vapours and hot objects: ICD-9 codes E958.1, E958.2; ICD-10 codes X76, X77
* self-inflicted injury by cutting and piecing instruments: ICD-9 code E956; ICD-10 codes X78, X79
* self-inflicted injury by jumping from high places: ICD-9 code E957, ICD-10 code X80
* self-inflicted injury by jumping or lying before a moving object: ICD-9 code E958.0, ICD-10 code X81
* self-inflicted injury by crashing of motor vehicle: ICD-9 code E958.5, ICD-10-CA code X82
* self-inflicted injury by other and unspecified means: ICD-9 codes E958.3, E958.4, E958.6-E958.9; ICD-10 codes X83, X84
* late effects of self-inflicted injury: ICD-9 code E959

To the above we can add late effects of self-inflicted injury in ICD-10: Y87.0, and remove the accidental poisonings!

## Compiling equivalent codes

```{r}
table_suicide_method_equivalence <- tribble(
  ~method, ~codes_intentional, ~codes_undetermined_intent,
  "poisoning", "E950-E952, X60-X69", "E980-E982, Y10-Y19", 
  "hanging, strangulation and suffocation", "E953, X70", "E983, Y20",
  "drowning", "E954, X71", "E984, Y21",
  "firearms and explosives", "E955, X72-X75", "E985, Y22-Y25",
  "smoke, fire, flames, steam, hot vapours and hot objects", "E958.1, E958.2, X76, X77", "E988.1, E988.2, Y26, Y27",
  "cutting and piercing instruments or blunt objects", "E956, X78, X79", "E986, Y28, Y29",
  "jumping from high places", "E957, X80", "E987, Y30",
  "jumping or lying before a moving object", "E958.0, X81", "E988.0, Y31",
  "crashing of motor vehicle", "E958.5, X82", "E988.5, Y32",
  "other and unspecified means", "E958.3, E958.4, E958.6, E958.7, E958.9, X83, X84", "E988.3, E988.4, E988.6, E988.7, E988.9, Y33, Y34",
  "sequelae", "E959, Y87.0", "E989, Y87.2"
)

if(!dir.exists("./table_suicide_methods")) dir.create("./table_suicide_methods")

write_csv(x = table_suicide_method_equivalence, path = "./table_suicide_methods/table_suicide_method_equivalence.csv")

convert_range_to_csv <- function(range_maybe_containing_dash) {
  if(!str_detect(range_maybe_containing_dash, pattern = "-")) return(range_maybe_containing_dash)
  first_char <- str_sub(range_maybe_containing_dash,1,1)
  numeric_range <-
    str_split(range_maybe_containing_dash, pattern = "-") %>% 
    unlist %>% 
    parse_number %>%
    {seq(from=.[1], to=.[2])}
  comma_separated_values <-
    paste0(first_char,numeric_range,collapse=",")
  return(comma_separated_values)
}

long_format_table_suicide_method_equivalence <-
  table_suicide_method_equivalence %>%
  pivot_longer(cols=c(codes_intentional,codes_undetermined_intent), names_to="intent", values_to="codes") %>%
  separate_rows(codes, sep=", ") %>%
  group_by(1:nrow(.)) %>%
  mutate(codes = convert_range_to_csv(codes)) %>%
  ungroup %>%
  separate_rows(codes, sep=",") %>%
  select(method, intent, code=codes) %>%
  mutate(
    intent = str_replace(string = intent, pattern = "codes\\_", replacement = ""),
    icd_version = if_else(condition = str_starts(string = code, pattern = "E"), true = 9, false = 10),
    code_decimal = code,
    code = str_replace(string = code, pattern = "\\.", replacement = "")
    ) %>%
  distinct

write_csv(x = long_format_table_suicide_method_equivalence, path = "./table_suicide_methods/table_suicide_method_equivalence_long.csv")
```


# Manual mapping

* ICD-10 X60 Intentional self-poisoning by and exposure to nonopioid analgesics, antipyretics and antirheumatics
* ICD-9 E950.0 Suicide and self-inflicted poisoning by analgesics, antipyretics, and antirheumatics


* X61 Intentional self-poisoning by and exposure to antiepileptic, sedative-hypnotic, antiparkinsonism and psychotropic drugs, not elsewhere classified
* E950.1 Suicide and self-inflicted poisoning by barbiturates
* E950.2 Suicide and self-inflicted poisoning by other sedatives and hypnotics

* X62 Intentional self-poisoning by and exposure to narcotics and psychodysleptics [hallucinogens], not elsewhere classified
* E950.3 Suicide and self-inflicted poisoning by tranquilizers and other psychotropic agents

* X63 Intentional self-poisoning by and exposure to other drugs acting on the autonomic nervous system
* ???

This is where it gets tricky - it's unclear which drugs in ICD-9 fall into "Other" and which don't!
