---
title: Extract & cross-map ICD codes from MVR papers and Schnitzer et al 2011
author: "Jan Savinc"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
  toc: true
toc_float: true
code_folding: hide
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The aim of this document is to produce an updated list of codes suggestive of maltreatment [@schnitzer_2011] - these were originally compiled for ICD-9-CM, and as part of the CHASe study [@dougall_2020] I translated them to ICD-9 and ICD-10, but for its use with US data, an ICD-10-CM version is needed.

# Libraries

We'll use `tidyverse` for data processing, and `icd` for its convenient functions for checking if codes exist in the ICD catalog. `knitr` is useful for printing nice tables in this document. `readxl` is for reading excel files!

```{r}
library(tidyverse)
library(icd)
library(knitr)
library(readxl)
```

## `icd` settings

`icd` will download data to a pre-defined location - we'll change that to a subfolder of this project.

```{r}
if (!dir.exists("./icd_data")) {
  dir.create("./icd_data")
}
set_icd_data_dir(path = "./icd_data")  # tell icd to download data here
# download_all_icd_data()  # tell icd to download everything
get_icd10cm2018()  # TODO: work out how this might work!
# get_icd10who2016()
```


## General Equivalence Mappings for the CM versions of ICD

The latest GEM files were downloaded from the [CDC website](ftp://ftp.cdc.gov/pub/Health_Statistics/NCHS/Publications/ICD10CM/2018/Dxgem_2018.zip). The 2018 release was used because it was the most recent at the time of writing.

```{r}
gem_icd9cm <- 
  read_fwf("./raw/2018_I9gem.txt", 
           fwf_cols(
             source = c(1, 5), 
             target = c(7, 13), 
             approximate=c(15,15), 
             no_map=c(16,16), 
             combination=c(17,17), 
             scenario=c(18,18), 
             choice_list=c(19,19) 
             )
           )

gem_icd10cm <- 
  read_fwf("./raw/2018_I10gem.txt", 
           fwf_cols(
             source = c(1, 7), 
             target = c(9, 13), 
             approximate=c(15,15), 
             no_map=c(16,16),  
             combination=c(17,17), 
             scenario=c(18,18), 
             choice_list=c(19,19) 
             )
           )

## for our purposes the distinction between forward & backward maps isn't relevant, so we can combine the two to simplify finding matches
gem_combined <- 
  bind_rows(
    gem_icd9cm %>% rename(icd9cm=source, icd10cm=target) %>% mutate(direction = "f"),
    gem_icd10cm %>% rename(icd9cm=target, icd10cm=source) %>% mutate(direction = "b")
  )
```

## Previously translated ICD-10 codes

```{r}
previously_translated_codes_file <- "./processed_ICD_codes/Schnitzer_et_al_2011_ICD_crossmapping_for_appendix_with_codes_and_descriptions.xlsx"

codes_translated <-
  map(
    .x = excel_sheets(previously_translated_codes_file),
    .f = ~read_excel(path = previously_translated_codes_file, sheet = .x)
  ) %>% set_names(x = ., nm = excel_sheets(previously_translated_codes_file))
```

# First attempt - load codes as-is

```{r}
codes_translated$inclusions %>%
  select(condition, icd10, comment_icd10) %>%
  mutate(
    icd10who = as.icd10who(icd10)#,
    # icd10cm = explain_code(icd10who)
    )

icd::explain_code("614.9")
icd::explain_code("N73")


icd::explain_code("9224")
icd::explain_code("S302")

icd::explain_code("V715")
icd::explain_code("Z044")

## observation for alleged abuse/neglext
icd::explain_code("V7181")
icd::explain_code("Z047")
icd::explain_code("Z0471")
icd::explain_code("Z047")
gem_combined %>% filter(str_detect(icd9cm,"V7181"))

## retinal haemorrhage
icd::explain_code("36281")
gem_combined %>% filter(str_detect(icd9cm,"36281"))
icd::explain_code("H356")

## rib fractures
icd::explain_code("8070")
icd::explain_code("8071")
icd::explain_code("S223") # one rib
icd::explain_code("S224") # multiple ribs

icd::explain_code("811")
icd::explain_code("S421")

## traumatic subdural hemorrhage
## with or without open wound is distinguished
icd::explain_code("8522")
icd::explain_code("8523")
icd::explain_code("S065")
icd::explain_code("S06893")

## stomach injury originally specified as injury with open wound but not in description so i think we can translate to more general code
icd::explain_code("8631")
icd::explain_code("S363")

## assault
icd::explain_code("E965")
icd::explain_code("X93")
icd::explain_code("X94")
icd::explain_code("X95")
icd::explain_code("X96")
icd::explain_code("E966")
icd::explain_code("X99")
icd::explain_code("E9682")
icd::explain_code("Y00")
## assault NOS
icd::explain_code("E9689")
icd::explain_code("Y09")

# TODO: this needs looking into!
## undet intent other means
icd::explain_code("E988")
icd::explain_code("Y26")

# TODO: this is also complicated - icd-10 splits vertebrae across body systems
# vertebral fracture
icd::explain_code("805")
icd::explain_code("S12")
# S12.0,S12.1,S12.2,S12.7,S12.9,S22.0,S22.1,S32.0,S32.7
icd::explain_code("S121")


```


