---
title: "Extracting ICD-9 codes from CDC materials"
author: "Jan Savinc"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NOTE: as of 24 April 2019 this document is abandoned!

# ICD-9 from GMWKI tables

TODO: clkean up text, explain
TODO: time the operations and move to separate doc to avoid re-computing everything again


The latest table from 1998 was used from:

https://www.cdc.gov/nchs/nvss/mortality/historical_gmwki.htm

Pages 1-110 contain listings

import using pdftools::pdf_text
creates a character vector with 1710 elements
first 110 elements correspond to pages 1-110

```{r}
ICD9_codes_gmwki_raw <-
  pdftools::pdf_text("../ICD codes/gmwki_98.pdf") %>%
  .[1:110] %>%
  str_split(pattern="\\r\\n") %>%
  unlist %>%
  tibble(raw=.) %>%
  filter(raw!="") %>%  # remove empty rows
  slice(-(1:8)) #%>%  # first 8 rows are introductory text and can be removed

ICD9_codes_gmwki_raw %>%
  filter(str_detect(raw, pattern = "^\\s{2}"))
```

# ICD-9-CM & ICD-10-CM General Equivalence Mapping (GEM)

The General Equivalence Mapping (GEM) is a set of two reference files that show the mapping from ICD-9-CM codes to ICD-10-CM codes and vice versa, where appropriate. We can use these to generate the cross-walk needed for the ICD-9-CM codes identified by Schnitzer et al., 2004 & 2011.

The ICD-10-CM codes identified this way will then need to be mapped further to WHO ICD-10 codes.

## Importing GEM files

The latest GEM files were downloaded from TODO: source

```{r}
gem_icd9cm <- 
  read_fwf("./raw/2018_I9gem.txt", 
           fwf_cols(
             source = c(1, 5), 
             target = c(7, 13), 
             approximate=c(15,15), 
             no_map=c(16,16), 
             combination=c(17,17), 
             scenario=c(18,18), 
             choice_list=c(19,19) 
             )
           )

gem_icd10cm <- 
  read_fwf("./raw/2018_I10gem.txt", 
           fwf_cols(
             source = c(1, 7), 
             target = c(9, 13), 
             approximate=c(15,15), 
             no_map=c(16,16), 
             combination=c(17,17), 
             scenario=c(18,18), 
             choice_list=c(19,19) 
             )
           )
```

## Doing the cross-walk

There are two directions of mapping: a forward mapping, from ICD-9 to ICD-10, and a backward mapping, from ICD-10 to ICD-9. Their semantics are slightly different, but irrelevant for our use - we are concerned with maximising sensitivity.

For our purposes, we will combine candidate codes from both the forward and backward mapping, and manually review them for appropriateness.

The mapping consists of: 

* single entry equivalents, where there is a one-to-one mapping (or multiple one-to-one mappings) for a given code. In these cases, the source code is sufficiently mapped by a single code (or by any of a range of single codes that are sufficiently similar in their meaning).
* combination entries, where there is a one-to-many mapping (or multiple one-to-many mappings) for a given code. In these cases, multiple target codes need to co-occur to sufficiently cover the meaning implied in the source code.

For our purposes, the above distinction is too complex, and any of the codes involved in any of the mappings will be used instead. For example, consider the ICD-9 code *E904.1, Accident due to lack of food* and its potential mappings:

Forward mapping: `r gem_icd9cm %>% filter(startsWith(source, prefix="E9041"))`
The no_map flag being 1 means there is no forward mapping available.

Backward mapping: `r gem_icd10cm %>% filter(startsWith(target, prefix="E9041"))`
A backward mapping to code *T730XXA Starvation, initial encounter* was found, but note that the combination flag is 1, which means this entry is part of a combination entry. The full entry is the following:

`r gem_icd10cm %>% filter(startsWith(source, prefix="T730XXA"))`

As you can see there are two entries in this combination entry. What this means is that to fully map the meaning of ICD-10-CM code *T730XXA Starvation, initial encounter*, both the codes *9942 Effects of hunger* AND *E904.1, Accident due to lack of food* would need to be included.

For our purposes, the (incomplete) backward mapping is sufficient, as we are not concerned with obtaining a fully equivalent mapping.


```{r}
#TODO: migrate to other doc

unmapped_icd9cm_codes_schnitzer_2004 <-
  schnitzer_2004_not_included_in_agi_2010$icd9code

unmapped_icd9cm_codes_schnitzer_2011 <-
  schnitzer_2011_not_included_in_agi_2010$icd9code

unmapped_icd9cm_codes_schnitzer_2004_nodecimal <- 
  unmapped_icd9cm_codes_schnitzer_2004 %>%
  gsub(., pattern="\\.", replacement = "")  # remove decimal

unmapped_icd9cm_codes_schnitzer_2011_nodecimal <- 
  unmapped_icd9cm_codes_schnitzer_2011 %>%
  gsub(., pattern="\\.", replacement = "")  # remove decimal



gem_icd9cm %>%
  filter(str_detect(source, pattern=paste(add_caret(unmapped_icd9cm_codes_schnitzer_2004_nodecimal),collapse = "|"))) %>%
  mutate(
    source_descr = sapply(source,function(x) icd::explain_code(x)),
    target_descr = sapply(target,function(x) icd::explain_code(x)),
    direction="forward"
    )

gem_icd10cm %>%
  filter(str_detect(target, pattern=paste(add_caret(unmapped_icd9cm_codes_schnitzer_2004_nodecimal),collapse = "|"))) %>%
  mutate(
    source_descr = sapply(source,function(x) icd::explain_code(x)),
    target_descr = sapply(target,function(x) icd::explain_code(x)),
    direction = "backward"
    )

map_forward_schnitzer_2011 <-
  gem_icd9cm %>%
  filter(str_detect(source, pattern=paste(add_caret(unmapped_icd9cm_codes_schnitzer_2011_nodecimal),collapse = "|"))) %>%
  mutate(
    source_descr = sapply(source,function(x) icd::explain_code(x)),
    target_descr = sapply(target,function(x) icd::explain_code(x)),
    direction="forward"
    )

map_backward_schnitzer_2011 <-
  gem_icd10cm %>%
  filter(str_detect(target, pattern=paste(add_caret(unmapped_icd9cm_codes_schnitzer_2011_nodecimal),collapse = "|"))) %>%
  mutate(
    source_descr = sapply(source,function(x) icd::explain_code(x)),
    target_descr = sapply(target,function(x) icd::explain_code(x)),
    direction="backward"
    )

#TODO: better way of doing this is to assemble (lapply?) the rows of GEM matching each individual code - that way we can also work out which code was being used as the filter
map_backward_schnitzer_2011_alt <-
  lapply(
    unmapped_icd9cm_codes_schnitzer_2011_nodecimal,
    FUN = function(x) {
      gem_icd10cm %>%
        filter(str_starts(target, pattern=x)) %>%
        mutate(
          schnitzer_code = x,
          schnitzer_descr = icd::explain_code(x),
          source_descr = sapply(source,function(x) icd::explain_code(x)),
          target_descr = sapply(target,function(x) icd::explain_code(x)),
          direction="backward"
        )
    } 
  ) %>%
  bind_rows

```

```{r}
save.image(file="./workspaces/codes_from_cdc_materials.RData")
```

