---
title: "Extracting Specialty codes pre-1997"
author: "Jan Savinc"
date: "4 November 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(tidyverse)
library(rvest)
```

# Introduction

The coding dictionary prior to 1997 is spread across the ISD website.

The dictionary for the hospital Specialty codes is at [this ISD website](https://www.ndc.scot.nhs.uk/Dictionary-A-Z/Definitions/index.asp?Search=S&ID=753&Title=SMR1/01%20Specialty%20Codes%20from%201961-96%20(SMR%20METADATA))

## Scraping html

Using *rvest*, we scrape the html code from the website and extract the table

```{r}
target_url <- "https://www.ndc.scot.nhs.uk/Dictionary-A-Z/Definitions/index.asp?Search=S&ID=753&Title=SMR1/01%20Specialty%20Codes%20from%201961-96%20(SMR%20METADATA)"

website <- read_html(target_url)
```

## Extracting the table

The first table on the website contains the coding:

```{r}
target_table_raw <- website %>% html_table(header = TRUE, trim = TRUE) %>% .[[1]]
```

## Cleanign up the table

To clean up the table, we perform the following steps:

* remove the entries before *80-81*
* replace all *--* entries with *NA*
* the 96 & 97 entries represent the new code, so we merge them
* we take the unique code across the years and define that as the old code

The entries for Clinical Chemistry, Palliative Medicine and Surgical Podiatry didn;t have a code prior to 1997 and are missing from the resulting table

```{r}
target_table <-
  target_table_raw %>%
  select(-one_of(c("61-66", "67-74", "75-77", "78-79"))) %>%
  rename(old_name="Old Specialty Name", new_name="Coppish Specialty Names") %>%
  mutate_all(~if_else(.=="--",NA_character_,.)) %>%
  mutate(
    new_code = case_when(
      !is.na(`97`) ~ `97`,
      !is.na(`96`) ~ `96`,
      TRUE ~ NA_character_
    )
  ) %>%
  select(-one_of(c("96","97"))) %>%
  gather(-new_name,-old_name,-new_code, key="year", value="value") %>%
  filter(!is.na(value)) %>%
  group_by(old_name) %>%
  mutate(old_code = paste(unique(value), collapse = ",")) %>%
  ungroup %>%
  select(-year, -value) %>%
  distinct
```

# Checking table

```{r}
kable(target_table, caption="Cleaned up Specialty code table")
```

Some old codes (e.g. *12* for Oral surgery & dental surgery) now map onto multiple codes (Oral surgery & Oral medicine, respectively).

# Saving resulting table

```{r}
write_csv(target_table, path = "./specialty_codes/old_specialty_codes_table.csv")
```


