---
title: "Format regression tables exported from Safe Haven for results paper"
author: "Jan Savinc"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries required

```{r}
library(tidyverse)  # for data manipulation
library(openxlsx)  # for writing excel files
library(readxl)  # for reading excel files
```

# Read tables

```{r}
order_icd_chapters <-
  c(
    "Infectious and parasitic diseases",
    "Neoplasms",
    "Endocrine, nutritional and metabolic diseases, and immunity disorders",
    "Diseases of the blood and blood-forming organs",
    "Diseases of the nervous system and the sense organs",
    "Mental disorders",
    "Diseases of the circulatory system",
    "Diseases of the respiratory system",
    "Diseases of the digestive system",
    "Diseases of the genitourinary system",
    "Complications of pregnancy, childbirth, and the puerperium",
    "Diseases of the skin and subcutaneous tissue",
    "Diseases of the musculoskeletal system and connective tissue",
    "Congenital anomalies",
    "Certain conditions originating in the perinatal period",
    "Symptoms, signs, and ill-defined conditions",
    "Injury and poisoning",
    "Supplementary classification of factors influencing health status and contact with health services"
  )

regression_adversities_female <- read_excel(path = "../Safe Haven Exports/2021-06-16 regressions/regressions/01_adversities_and_geography/table_regression_results_female_adversities.xlsx", sheet = "multiple reg., simplified")

regression_adversities_male <- read_excel(path = "../Safe Haven Exports/2021-06-16 regressions/regressions/01_adversities_and_geography/table_regression_results_male_adversities.xlsx", sheet = "multiple reg., simplified")

regression_subcategorised_adversities_female <- read_excel(path = "../Safe Haven Exports/2021-06-16 regressions/regressions/02_adversities_subcategorised/table_regression_results_female_adversities.xlsx", sheet = "multiple reg., simplified")

regression_subcategorised_adversities_male <- read_excel(path = "../Safe Haven Exports/2021-06-16 regressions/regressions/02_adversities_subcategorised/table_regression_results_male_adversities.xlsx", sheet = "multiple reg., simplified")
```

## Variable names

```{r}
table_variable_names <-
  tribble(
    ~short_name, ~long_name,
    "n_episodes_mvr_under_18_factor", "MVR episodes, aged <18",
    "n_episodes_ccs_under_18_factor", "MH episodes, aged <18",
    # "n_episodes_schnitzer_without_caries_under_10_factor", "Episodes with codes suggestive of maltreatment or neglect (Schnitzer et al., 2011), excluding dental caries, aged <10",
    "n_episodes_schnitzer_without_caries_under_10_factor", "Episodes with codes suggestive of maltreatment or neglect, excluding dental caries, aged <10",
    "n_episodes_poisoning_under_18_factor", "Poisoning episodes, aged <18",
    "mother_died_before_age_18", "Maternal death, aged <18",
    "care_experienced_or_no_fixed_abode_before_age_18", "Episodes indicating care experience or no fixed abode, aged <18",
    "n_episodes_mvr_assault_maltreatment_under_18_factor", "MVR episodes: assault or maltreatment, aged <18",
    "n_episodes_mvr_social_or_undetermined_under_18_factor", "MVR episodes: adverse soc. circumstances or undetermined intent, aged <18",
    "n_episodes_ccs_self_harm_under_18_factor", "MH episodes: self harm, aged <18",
    "n_episodes_ccs_alcohol_under_18_factor", "MH episodes: alcohol-related, aged <18",
    "n_episodes_ccs_not_self_harm_or_alcohol_under_18_factor", "MH episodes: neither self harm nor alcohol-related, aged <18"
  )
```


# Process tables


For the "simple" adversities table, we include all adversities including Schnitzer codes without dental caries, but not the geographical variables. These are models F2 and M2 in the outputs!

```{r}
regression_table_simple_adversities <-
  full_join(
    regression_adversities_female %>% select(variable, level, F2_or, F2_global_p),
    regression_adversities_male %>% select(variable, level, M2_or, M2_global_p),
    by = c("variable","level")
  ) %>%
  filter(!variable %in% c("Urban_rural_aggregated_4")) %>%
  mutate(  # rename variables to long form, descriptive names
    variable = table_variable_names$long_name[match(x = variable, table = table_variable_names$short_name)]
  ) %>%
  rename(
    Variable = variable,
    Level = level,
    "OR, Female" = F2_or,
    "p, Female" = F2_global_p,
    "OR, Male" = M2_or,
    "p, Male" = M2_global_p,
  )
```


For the "subcategorised" adversities table, we include all subcategoriesd adversities including Schnitzer codes without dental caries, but not the geographical variables. These are models F5 and M5 in the outputs!

```{r}
regression_table_subcategorised_adversities <-
  full_join(
    regression_subcategorised_adversities_female %>% select(variable, level, F5_or, F5_global_p),
    regression_subcategorised_adversities_male %>% select(variable, level, M5_or, M5_global_p),
    by = c("variable","level")
  ) %>%
  filter(!variable %in% c("Urban_rural_aggregated_4")) %>%
  mutate(  # rename variables to long form, descriptive names
    variable = table_variable_names$long_name[match(x = variable, table = table_variable_names$short_name)]
  ) %>%
  rename(
    Variable = variable,
    Level = level,
    "OR, Female" = F5_or,
    "p, Female" = F5_global_p,
    "OR, Male" = M5_or,
    "p, Male" = M5_global_p,
  )
```

# Save regression tables

```{r}
write.xlsx(x = regression_table_simple_adversities, file = "../Publications/Results paper 1/Table_multiple_regression_simple_from_r.xlsx")
write.xlsx(x = regression_table_subcategorised_adversities, file = "../Publications/Results paper 1/Table_multiple_regression_subcategorised_from_r.xlsx")
```


# OR plots

## Read data

For the regression tables, I opted to use the simplified tables i made in the safe haven; for OR plots, we need access to confidence interval estimates, so I'll load the data again, but from the more detailed outputs!

```{r}
detailed_regressions_simple_female <-
  read_excel(path = "../Safe Haven Exports/2021-06-16 regressions/regressions/01_adversities_and_geography/table_regression_results_female_adversities.xlsx", sheet = "multiple regression")

detailed_regressions_simple_male <-
  read_excel(path = "../Safe Haven Exports/2021-06-16 regressions/regressions/01_adversities_and_geography/table_regression_results_male_adversities.xlsx", sheet = "multiple regression")

detailed_regressions_subcategorised_female <-
  read_excel(path = "../Safe Haven Exports/2021-06-16 regressions/regressions/02_adversities_subcategorised/table_regression_results_female_adversities.xlsx", sheet = "multiple regression")

detailed_regressions_subcategorised_male <-
  read_excel(path = "../Safe Haven Exports/2021-06-16 regressions/regressions/02_adversities_subcategorised/table_regression_results_male_adversities.xlsx", sheet = "multiple regression")
```

## Process data

Again, we'll use the models with Schnitzer episodes included, so F2 & M2

```{r}
detailed_regressions_simple <-
  bind_rows(
    detailed_regressions_simple_female %>% select(variable, level, starts_with("F2")) %>%
    filter(!variable %in% c("Urban_rural_aggregated_4")) %>%
    rename_all(~str_remove_all(.x, pattern="F2\\_")) %>%
    mutate(sex = "Female")
    ,
    detailed_regressions_simple_male %>% select(variable, level, starts_with("M2")) %>%
    filter(!variable %in% c("Urban_rural_aggregated_4")) %>%
    rename_all(~str_remove_all(.x, pattern="M2\\_")) %>%
    mutate(sex = "Male")
  )

detailed_regressions_subcategorised <-
  bind_rows(
    detailed_regressions_subcategorised_female %>% select(variable, level, starts_with("F5")) %>%
    filter(!variable %in% c("Urban_rural_aggregated_4")) %>%
    rename_all(~str_remove_all(.x, pattern="F5\\_")) %>%
    mutate(sex = "Female")
    ,
    detailed_regressions_subcategorised_male %>% select(variable, level, starts_with("M5")) %>%
    filter(!variable %in% c("Urban_rural_aggregated_4")) %>%
    rename_all(~str_remove_all(.x, pattern="M5\\_")) %>%
    mutate(sex = "Male")
  )
```

## Make plots

```{r}
# TODO: should this include Intercept? Intercept tells us the baseline OR of cases to controls - over 0.1 (because ratio is less than 1:10!)

wrap_width = 50  # to control how wide the variable labels are spread

(
or_plot_simple_adversities <-
  detailed_regressions_simple %>%
  mutate(  # rename variables to long form, descriptive names
    variable = table_variable_names$long_name[match(x = variable, table = table_variable_names$short_name)],
    variable = if_else(level == "(Intercept)", level, variable),
    level = if_else(level %in% c("(Intercept)", "TRUE"), NA_character_, level),
    variable = if_else(is.na(level), str_wrap(variable, width = 30), paste(str_wrap(variable, width = 30), level, sep = ":\n")),
    variable = forcats::as_factor(variable) %>% fct_rev(),  # this retains the order!
  ) %>%
  ggplot(aes(y = variable, x = or, shape = sex)) +
  geom_point(position = position_dodge(width=0.5), size = 2.5) +
  geom_errorbar(aes(xmin = or_ci_low, xmax = or_ci_high), position = position_dodge(width=0.5), width = 0) +
  theme_minimal() +
  scale_x_log10(breaks = c(seq(from=0.1, to=0.9, by=0.1), 1:10, 15, 30), labels = c(as.character(seq(from=0.1, to=0.9, by=0.1)), 1:10, 15, 30)) +
  geom_vline(xintercept = 1, colour = "grey10", linetype = "dashed") +
  theme(
    legend.position = "top", 
    legend.title = element_blank(), 
    panel.grid.minor.x = element_blank()
    ) +
  labs(
    y = NULL,
    x = "OR"
  ) +
  NULL
  )


(
or_plot_subcategorised_adversities <-
  detailed_regressions_subcategorised %>%
  mutate(  # rename variables to long form, descriptive names
    variable = table_variable_names$long_name[match(x = variable, table = table_variable_names$short_name)],
    variable = if_else(level == "(Intercept)", level, variable),
    level = if_else(level %in% c("(Intercept)", "TRUE"), NA_character_, level),
    variable = if_else(is.na(level), str_wrap(variable, width = wrap_width), paste(str_wrap(variable, width = wrap_width), level, sep = ":\n")),
    variable = forcats::as_factor(variable) %>% fct_rev()  # this retains the order!
  ) %>%
  ggplot(aes(y = variable, x = or, shape = sex)) +
  geom_point(position = position_dodge(width=0.5), size = 2.5) +
  geom_errorbar(aes(xmin = or_ci_low, xmax = or_ci_high), position = position_dodge(width=0.5), width = 0) +
  theme_minimal() +
  scale_x_log10(breaks = c(seq(from=0.1, to=0.9, by=0.1), 1:10, 15, 30), labels = c(as.character(seq(from=0.1, to=0.9, by=0.1)), 1:10, 15, 30)) +
  geom_vline(xintercept = 1, colour = "grey10", linetype = "dashed") +
  theme(
    legend.position = "top", 
    legend.title = element_blank(), 
    panel.grid.minor.x = element_blank(),
    axis.text.y = element_text(size = 11),
    ) +
  labs(
    y = NULL,
    x = "Adjusted odds ratio"
  ) +
  NULL
  )
```

## Save plots

```{r}
## changed this on 3 Feb 2022 to save to OneDrive folder
plot_directory <- "C:/Users/40011625/OneDrive - Edinburgh Napier University/CHASE Publications/Results paper 1/"

or_plot_simple_adversities %>%
  ggsave(filename = file.path(plot_directory, "OR_plot_regression_simple.pdf"), device = cairo_pdf, dpi = 300, width = 12, height = 8)

or_plot_simple_adversities %>%
  ggsave(filename = file.path(plot_directory, "OR_plot_regression_simple.png"), dpi = 1200, width = 12, height = 8)

or_plot_subcategorised_adversities %>%
  ggsave(filename = file.path(plot_directory, "OR_plot_regression_subcategorised.pdf"), device = cairo_pdf, dpi = 300, width = 13, height = 8)

or_plot_subcategorised_adversities %>%
  ggsave(filename = file.path(plot_directory, "OR_plot_regression_subcategorised.png"), dpi = 1200, width = 13, height = 8)
```

