---
title: CHASe Hypothesis testing using data exported from Safe Haven
author: "Jan Savinc"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
  toc: true
toc_float: true
code_folding: hide
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r, warning=FALSE}
library(tidyverse)
library(knitr)  # for showing tables inside rmarkdown
library(readxl)  # read excel files
library(broom)
```

# Load data

```{r}
consort_data <- 
  read_csv(file = "../Safe Haven Exports/CONSORT_diagram_data.csv") %>%
  mutate(
    note = case_when(
      note == "Excluded: cases with no hospital records prior to death before age 18" &
        N > 1000 ~ "Excluded: controls with no hospital records prior to death before age 18",
      TRUE ~ note
    )
  )  # correction: instead of 'cases' it should be 'controls' at the entry for exclusions with no records prior to death or age 18 with N>1000

data_cohort_by_whether_they_had_records_prior_to_death <- read_excel(
  path = "X:/R1369/CSO FULL grant/Safe Haven Exports/Descriptive summary of cohort & individuals with any records/Descriptive_summary_cohort_by_whether_they_had_any_records_prior_to_death.xlsx",
  sheet = 1, 
  ) %>%
  slice(-13)  # remove footnote entry in spreadsheet

data_descriptives_carstairs_simd <- read_excel(
  path = "X:/R1369/CSO FULL grant/Safe Haven Exports/Exports 2020-10-07/Results_paper_1_tables/Descriptives_carstairs_and_simd.xlsx", sheet = 1
) %>% slice(-nrow(.))  # remove last row - it's the footnote

hypothesis_tests <- list()  # empty list for storing hypothesis tests!
```


## Hypothesis: the proportion of males is higher in the suicide group with hospital records than in the control group with hospital records


```{r}
table_cohort_sex_has_any_records_prior_to_death <-
  data_cohort_by_whether_they_had_records_prior_to_death %>%
  select(case, sex, has_no_records_prior_to_death, N)

## print the table first so we see the data we're working with
table_cohort_sex_has_any_records_prior_to_death %>%
  filter(sex!="both" & has_no_records_prior_to_death=="No") %>%
  group_by(case) %>%
  mutate(percent = scales::percent(N/sum(N), accuracy = 0.1)) %>%
  ungroup %>%
  kable(caption = "Number of individuals with records prior to death by sex & case/control.")

## the proportions look very similar
table_cohort_sex_has_any_records_prior_to_death %>%
  filter(sex!="both" & has_no_records_prior_to_death=="No") %>%
  uncount(N) %>%
  {table(.$case, .$sex)} %>%
  chisq.test()
```


## Hypothesis: the average age of death is different for suicide cases with no prior records to those with prior records

```{r}
mean_cases_with_records <- data_cohort_by_whether_they_had_records_prior_to_death %>% filter(case=="case" & sex=="both" & has_no_records_prior_to_death=="No") %>% .$mean_age
mean_cases_without_records <- data_cohort_by_whether_they_had_records_prior_to_death %>% filter(case=="case" & sex=="both" & has_no_records_prior_to_death=="Yes") %>% .$mean_age

sd_cases_with_records <- data_cohort_by_whether_they_had_records_prior_to_death %>% filter(case=="case" & sex=="both" & has_no_records_prior_to_death=="No") %>% .$sd_age
sd_cases_without_records <- data_cohort_by_whether_they_had_records_prior_to_death %>% filter(case=="case" & sex=="both" & has_no_records_prior_to_death=="Yes") %>% .$sd_age

n_cases_with_records <- data_cohort_by_whether_they_had_records_prior_to_death %>% filter(case=="case" & sex=="both" & has_no_records_prior_to_death=="No") %>% .$N
n_cases_without_records <- data_cohort_by_whether_they_had_records_prior_to_death %>% filter(case=="case" & sex=="both" & has_no_records_prior_to_death=="Yes") %>% .$N

## convert table to a format of one entry per row
age_at_death_tbl <-
  data_cohort_by_whether_they_had_records_prior_to_death %>%
  filter(!is.na(mean_age)) %>%
  select(case, sex, has_no_records_prior_to_death, N, mean_age, sd_age) %>%
  mutate(has_no_records_prior_to_death = if_else(has_no_records_prior_to_death=="Yes", "no_records", "had_records")) %>%
  pivot_wider(values_from = c(N, mean_age, sd_age), names_from = has_no_records_prior_to_death)

## t test using statistics rather than raw data
hypothesis_tests$t_test_age_difference_cases_with_and_without_records <-
  age_at_death_tbl %>%
  group_by(case, sex) %>%
  summarise(
    tidy(BSDA::tsum.test(
      mean.x = mean_age_had_records,
      s.x = sd_age_had_records,
      n.x = N_had_records,
      mean.y = mean_age_no_records,
      s.y = sd_age_no_records,
      n.y = N_no_records,
      var.equal = FALSE
    )),
    .groups = "drop"
  ) %>%
  mutate(
    difference = estimate1 - estimate2,
    meaning = "1 = had records, 2 = no records"
    ) %>%
  relocate(difference, .before=statistic)

hypothesis_tests$t_test_age_difference_cases_with_and_without_records %>%
  kable(caption = "t-test of age difference between cases with records and cases without records", format.args = list(nsmall=4))
```

## Hypotheses: gender differences between characteristics of cases & controls, separately for individuals with records or no records

```{r}
# quantities_to_test_with_t_test <- c("age","carstairs","urban_rural")
# quantities_to_test_with_chisq_test <- c("proportion_not_in_work")

long_data_cohort_by_whether_they_had_records_prior_to_death_for_t_tests <-
  bind_rows(
  ## AGE  
    data_cohort_by_whether_they_had_records_prior_to_death %>%  # this uses deciles for carstairs so we'll take the newer quintile carstairs data from different table
    # filter(sex != "both") %>%
    mutate(n_age = N, n_carstairs = N, n_urban_rural = N) %>%
    select(-matches("median|iqr|q1|q3|proportion"), -N) %>%
    pivot_longer(cols = matches("age|carstairs|urban_rural"), names_to = c("statistic","variable"), values_to = "value", names_pattern = "(mean|sd|n)_(.*)") %>%
    filter(variable=="age" & case=="case")  # only keep age, which only makes sense for cases
    ,
    ## CARSTAIRS & SIMD
    data_descriptives_carstairs_simd %>% 
    # filter(measure=="quintile") %>%
    mutate(variable=paste0(tolower(geographic_indicator),"_",measure)) %>%
    rename_all(~tolower(.)) %>%
    select(-matches("median|iqr|q1|q3|proportion|geographic_indicator|^measure")) %>%
    pivot_longer(cols=c(n,mean,sd), names_to="statistic",values_to="value")
  )

hypothesis_tests$t_tests_sex_differences <-
  long_data_cohort_by_whether_they_had_records_prior_to_death_for_t_tests %>%
  filter(sex!="both") %>%
  pivot_wider(names_from = c("sex","statistic"), values_from = "value") %>%
  group_by(case, has_no_records_prior_to_death, variable) %>%
  summarise(
    BSDA::tsum.test(
      mean.x = female_mean,
      s.x = female_sd,
      n.x = female_n,
      mean.y = male_mean,
      s.y = male_sd,
      n.y = male_n,
      var.equal = FALSE
      ) %>% tidy,
    difference = estimate1-estimate2,
    meaning = "female - male (1=female)",
    .groups = "drop"
    ) %>%
  relocate(meaning, difference, .after = variable)

hypothesis_tests$t_tests_cohort_differences <-
  long_data_cohort_by_whether_they_had_records_prior_to_death_for_t_tests %>%
  filter(variable!="age") %>%  # can't compare age, not sensible for controls
  filter(has_no_records_prior_to_death != "Both") %>%  # without splitting by had records prior to death, all descriptives are the same! (cases matched to control on geography)
  pivot_wider(names_from = c("case","statistic"), values_from = "value") %>%
  group_by(sex, has_no_records_prior_to_death, variable) %>%
  summarise(
    BSDA::tsum.test(
      mean.x = case_mean,
      s.x = case_sd,
      n.x = case_n,
      mean.y = control_mean,
      s.y = control_sd,
      n.y = control_n,
      var.equal = FALSE
      ) %>% tidy,
    difference = estimate1-estimate2,
    meaning = "case - control (1=case)",
    .groups = "drop"
    ) %>%
  relocate(meaning, difference, .after = variable)

## Proportion tests of not in work
hypothesis_tests$prop_tests_sex_differences <-
  data_cohort_by_whether_they_had_records_prior_to_death %>%
  filter(sex != "both") %>%
  filter(!is.na(proportion_not_in_work)) %>%
  select(case, sex, has_no_records_prior_to_death, N, proportion_not_in_work) %>%
  mutate(n_success = round(N * proportion_not_in_work), n_fail = N-n_success) %>%
  select(-N, -proportion_not_in_work) %>%
  pivot_longer(cols=c(n_success,n_fail), names_to="outcome", values_to="n", names_pattern="n_(.*)") %>%
  uncount(n) %>%
  group_by(case, sex, has_no_records_prior_to_death) %>%
  summarise(
    prop.test(table(outcome)) %>% tidy,
    .groups = "drop"
  )
```

### Display results

```{r}
hypothesis_tests$t_tests_sex_differences %>%
  filter(p.value < .05) %>%
  kable(caption = "t-tests of sex difference at p < .05", format.args = list(nsmall=4))

hypothesis_tests$t_tests_cohort_differences %>%
  filter(p.value < .05) %>%
  kable(caption = "t-tests of cohorty (case - control) difference at p < .05", format.args = list(nsmall=4))

hypothesis_tests$prop_tests_sex_differences %>%
  filter(p.value < .05) %>%
  kable(caption = "Fisher's z-tests of sex difference at p < .05", format.args = list(nsmall=4))
```

