---
title: "Merge exported Scopus & Web of Science records"
author: "Jan Savinc"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

For the ICD code literature review, I performed a search on Scopus & Web of Science (WoS), resulting in 300-400 records each. To avoid duplicating the work of finding relevant records, this script will find records that appear in both sets and produce a merged set.

```{r}
library(tidyverse)
library(openxlsx)
```

# Loading the files

The Scopus file is in a much neater format, so we'll use this as the base file.

We rename the WoS columns to match Scopus first.

```{r}
file_scopus <- read.csv("./lit_review_merge_records/scopus.csv", stringsAsFactors = FALSE, encoding="UTF-8") %>%
  rename(Authors=X.U.FEFF.Authors)
file_wos <- 
  read.csv("./lit_review_merge_records/savedrecs.tsv", sep="\t", stringsAsFactors = FALSE, strip.white = TRUE, fileEncoding = "UTF-16LE", row.names=NULL) %>%
  select(  # rename the relevant columns to Scopus columns, keeping only ones that are named below - the rest is a mess and not relevant
    Authors=GP,
    Title=CA,
    Year=PD,
    Source.title=TI,
    DOI=AR,
    Abstract=ID
  ) %>%
  mutate(Source="Web of Science")
```

The Scopus data has `r nrow(file_scopus)` records, and the Web of Science has `r nrow(file_wos)` records.

# Merging the records

Now we go through the records in the WoS file and check whether they are in the Scopus data. If they aren't they'll be added on to the final file.

## First pass: remove identical titles

```{r}
dataToBeAddedToScopus <- file_wos %>%
  filter(!Title %in% file_scopus$Title)
```

We now have `r nrow(dataToBeAddedToScopus)` records left in WoS.

## Second pass: algorithm

This is based on a simple algorithm copied from: https://github.com/christopherkullenberg/WoScop

1. titles are converted to lower case
2. all punctuation and whitespace are removed
3. the first X characters of each thus geenrated string will be matched

```{r}
file_scopus$uniqueTitle <-
   gsub(file_scopus$Title, pattern="[^A-Za-z0-9]+", replacement="") %>% tolower

dataToBeAddedToScopus$uniqueTitle <-
   gsub(dataToBeAddedToScopus$Title, pattern="[^A-Za-z0-9]+", replacement="") %>% tolower

dataToBeAddedToScopus <-
  dataToBeAddedToScopus %>%
  filter(!uniqueTitle %in% file_scopus$uniqueTitle)
```

After applying the algorithm and removing the matches, we are left with `r nrow(dataToBeAddedToScopus)` records.

## Third pass: approximate matching

Now we use approximate (fuzzy) matching based on the algorithmically derived titles.

```{r}
foundFuzzyMatches <- tibble(
  wos = dataToBeAddedToScopus$Title,
  fuzzyMatches = sapply(
    X = dataToBeAddedToScopus$uniqueTitle,
    FUN = function(x) paste(file_scopus$Title[agrep(file_scopus$uniqueTitle,pattern = x)],collapse = "; "), USE.NAMES = FALSE
  )
) %>%
  filter(!is.na(fuzzyMatches) & fuzzyMatches!="")

foundFuzzyMatches
```

Only the three approxiamte matches displayed above were found, and on manual review the final two pairs are actual matches, but not the first pair.

```{r}
dataToBeAddedToScopus <-
  dataToBeAddedToScopus %>%
  filter(!Title %in% foundFuzzyMatches$wos[2:3])
```

This leaves us with `r nrow(dataToBeAddedToScopus)` most likely unique WoS records. They can now be merged with the Scopus records for manual review.

# Merging data

```{r}
mergedData <-
  bind_rows(
    file_scopus,
    dataToBeAddedToScopus
  ) %>%
  select(-uniqueTitle) %>%  # no longer needed
  mutate(
    DOI_link = ifelse(
      !is.na(DOI) & DOI!="",
      paste0("https://doi.org/",DOI),
      NA
    )
  )

class(mergedData$Link) <- "hyperlink"
class(mergedData$DOI_link) <- "hyperlink"
```

We have a total of (most likely unique) `r nrow(mergedData)` records now, and we save them to an excel file for further processing.

# Saving merged data

The merged data will be save to file *merged_records_Scopus_WoS.xlsx*

```{r}
write.xlsx(x = list("merged Scopus,WoS"=mergedData), file="./lit_review_merge_records/merged_records_Scopus_WoS.xlsx")
```



