---
title: Match inclusion & exclusion ICD codes from Schnitzer et al. (2011) translated to ICD-10-CM to a sample dataset 
author: "Jan Savinc"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
  toc: true
  toc_float: true
  code_folding: hide
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

The aim of this document is to serve as an example of the use of codes suggestive of maltreatment [@schnitzer_2011] translated to ICD-10-CM for use with medical records in the US. The algorithm used for matching inclusion & exclusion codes is shown.

The original codes [@schnitzer_2011] were originally compiled for ICD-9-CM, and as part of the CHASe study [@dougall_2020] I translated them to ICD-9 and ICD-10; I later translated them to ICD-10-CM also!


# Libraries

```{r}
library(tidyverse)  # for tidyverse-style code
library(knitr)  # for printing tables
library(fuzzyjoin)  # for regular expression joins
library(tictoc)  # for timing
```

# Load data

```{r}
tic()  # start timing

admissions <- read_csv("../Collaborations/translated_to_icd10cm/Sample Set of Simulated 3 Yrs of Encounters.csv") %>% 
  mutate(episode_id = 1:n()) %>%  # add episode id
  distinct

criteria <- list(
  inclusions = read_csv("../Collaborations/translated_to_icd10cm/codes_translated_to_icd10cm/codes_schnitzer2011_icd10cm_inclusions.csv"),
  all_exclusions = read_csv("../Collaborations/translated_to_icd10cm/codes_translated_to_icd10cm/codes_schnitzer2011_icd10cm_all_exclusions.csv"),
  additional_requirements = read_csv("../Collaborations/translated_to_icd10cm/codes_translated_to_icd10cm/codes_schnitzer2011_icd10cm_additional_requirements.csv")
)
```

# Convert data to long format

Long format in this case means having one diagnosis per row - this simplifies processing! The episode ID defined above will enable finding inclusion & exclusion criteria for codes within the context of the same episode.

```{r}
# We convert the admissions into a format where each admission has a unique identifier, and each row represents one diagnosis code
# admissions_in_long_format <- NULL  # whatever processing needs doing to get to that stage
admissions_in_long_format <- 
  admissions %>% 
  pivot_longer(
    cols = starts_with("diagnosis"),
    names_to = "source",
    values_to = "code",
    values_drop_na = TRUE
  )
```


# Inclusion / exclusion matching algorithm

Each episode can have one or more associated diagnosis codes.

The inclusion and exclusion codes comprise the following:

* Inclusions
  - Additional requirements - these are codes that need to be present in the same episode in addition to the 2nd hand tobacco smoke diagnosis to be included
* Exclusions - these are defined for certain inclusion codes, so we only include the relevant codes if none of the other codes in the same episode
  - Malnutrition exclusions - these are specified separately for convenience in translation only so they are to be merged with the other exclusion

The codes are specified in a human-readable format (in columns `inclusion_code`, `exclusion_code`, or `additional_inclusion_code`), as well as a `regex_prefix` variable, which is a *regular expression* designed to capture the relevant sequence of characters. In most cases this is just a string prefaced by a caret (`^`), to denote the beginning of a string, e.g. `^J45`, but there are a few codes for undetermined intent poisonings where a more complex sequence is matched: to distinguish `T36` codes with undetermined intent (6th character being `4`: `^T36..4`) from accident, self-harm or assault codes (6rh character *not* being `4`: `^T36..[^4]`).

The algorithm outline is as follows:

1. Create a subset of codes that match any of the inclusion codes (including the age criterion)
2. Remove inclusions that don't match the additional requirements
  - Take subset of included codes that have additional requirements
  - Take episode ids of codes with additional requirements and take subset of episodes that match additional requirement codes
  - Keep intersect of episodes (episode ids) that have an inclusion code and also its associated additional requirement
3. From the subset of episodes with inclusion codes create a subset of codes that match any of the exclusions
4. Find intersect between episodes with inclusions and episodes with exclusions that match the inclusions (using episode id and inclusion id)
5. Keep subset of inclusion codes with the intersect from 4. removed

## Step 1: Create a subset of codes that match any of the inclusion codes (including the age criterion)

```{r}
admissions_with_inclusions <-
  admissions_in_long_format %>%
  select(id, episode_id, code, age) %>%
  regex_inner_join(
    criteria$inclusions, by = c("code"="regex_prefix")
  ) %>%
  filter(age<age_less_than)
```

## Step 2: Remove inclusions that don't match the additional requirements

```{r}
## what inclusion index was used for codes with additional requirements?
inclusion_index_with_additional_requirements <-
  criteria$additional_requirements %>%
  .$inclusion_index %>%
  unique

## take subset of includes codes that have additional requirements
admissions_with_inclusions_with_additional_requirements <-
  admissions_with_inclusions %>%
  filter(inclusion_index %in% inclusion_index_with_additional_requirements)

## take subset of episodes that contain codes with additional requirements
episodes_with_additional_requirements <-
  admissions_with_inclusions_with_additional_requirements %>%
  select(id, episode_id) %>%  # keep id & episode ids
  distinct %>%
  left_join(admissions_in_long_format, by = c("id", "episode_id")) %>%
  select(id, episode_id, code) %>%
  regex_inner_join(  # keep only episodes that have at least one additional requirement
    criteria$additional_requirements, by=c("code"="regex_prefix")
  )

admissions_with_additional_requirements_satisfied <-
  admissions_with_inclusions_with_additional_requirements %>%
  semi_join(episodes_with_additional_requirements, by = c("id","episode_id")) %>%  # keep admission with inclusion codes that have episode ids that were found to have any of the additional requirements also
  distinct

admissions_with_inclusion_but_additional_requirements_not_satisfied <-
  admissions_with_inclusions_with_additional_requirements %>%
  anti_join(episodes_with_additional_requirements, by = c("id","episode_id")) %>%  # keep admission with inclusion codes that have episode ids that were found to have any of the additional requirements also
  distinct
  
## define the narrower subset of episodes with inclusion codes- remove codes with additional requirements, and then add in those episodes with additional requirements where the additional requirements were satisfied
admissions_with_inclusions_after_sorting_additional_requirements <-
  admissions_with_inclusions %>%
  anti_join(
    admissions_with_inclusions_with_additional_requirements, by = c("id", "episode_id", "code")
  ) %>%
  bind_rows(
    .,
    admissions_with_additional_requirements_satisfied
  ) %>%
  distinct
```


## Step 3: From the subset of episodes with inclusion codes create a subset of codes that match any of the exclusions

```{r}
episodes_with_inclusions <-
  admissions_with_inclusions_after_sorting_additional_requirements %>%
  select(id, episode_id) %>%
  distinct

episodes_matching_any_exclusions <-
  episodes_with_inclusions %>%
  left_join(admissions_in_long_format, by = c("id", "episode_id")) %>%
  select(id, episode_id, code) %>%
  regex_inner_join(  # keep only episodes that have at least one additional requirement
    criteria$all_exclusions, by=c("code"="regex_prefix")
  )
```


## Step 4: Find intersect between episodes with inclusions and episodes with exclusions that match the inclusions (using episode id and inclusion id)

```{r}
intersect_between_inclusions_and_exclusions <-
  admissions_with_inclusions_after_sorting_additional_requirements %>%
  inner_join(
    episodes_matching_any_exclusions,
    by = c("id","episode_id","inclusion_index")
  )
```


## Step 5: Keep subset of inclusion codes with the intersect from 4. removed

```{r}
inclusion_codes_with_exclusions_removed <-
  admissions_with_inclusions_after_sorting_additional_requirements %>%
  anti_join(
    intersect_between_inclusions_and_exclusions,
    by = c("id","episode_id","inclusion_index")
  )
```

# Results of the process

The processing of the data took this long:

```{r}
toc()  # display how much time elapsed since starting timing!
```

A total of N=`r format(nrow(admissions), big.mark=",")` admissions were processed or N=`r format(nrow(admissions_in_long_format), big.mark=",")` of diagnosis codes, belonging to N=`r format(n_distinct(admissions$id),big.mark=",")` individuals.

A total of N=`r format(nrow(inclusion_codes_with_exclusions_removed), big.mark=",")` of codes suggestive of maltreatment or neglect were found, belonging to N=`r format(n_distinct(inclusion_codes_with_exclusions_removed$episode_id,inclusion_codes_with_exclusions_removed$id), big.mark=",")` episodes, or N=`r format(n_distinct(inclusion_codes_with_exclusions_removed$id), big.mark=",")` individuals.

## Breakdown of maltreatment categories according to Schnitzer et al., 2011

```{r}
inclusion_codes_with_exclusions_removed %>% 
  count(maltreatment_category) %>%
  mutate(percent = scales::percent(n/sum(n), accuracy = 0.1)) %>%
  add_case(maltreatment_category="TOTAL", n = sum(.$n), percent="100%")
```

## Number of episodes by individuals

```{r}
inclusion_codes_with_exclusions_removed %>%
  group_by(id) %>%
  summarise(n_episodes = n_distinct(episode_id), .groups = "drop") %>%
  complete(id = admissions$id, fill = list(n_episodes=0)) %>%  # add in all the ids from all admissions, adding 0 episodes for those not yet included
  count(n_episodes) %>%
  mutate(percent = scales::percent(n/sum(n), accuracy = 0.1))

## number of individuals with any episodes
inclusion_codes_with_exclusions_removed %>%
  group_by(id) %>%
  summarise(n_episodes = n_distinct(episode_id), .groups = "drop") %>%
  complete(id = admissions$id, fill = list(n_episodes=0)) %>%  # add in all the ids from all admissions, adding 0 episodes for those not yet included
  count(n_episodes > 0) %>%
  mutate(percent = scales::percent(n/sum(n), accuracy = 0.1))
```

# Tagging original datasets with inclusions

```{r}
admissions_tagged_with_inclusions <-
  admissions %>%
  mutate(
    meets_all_criteria = id %in% inclusion_codes_with_exclusions_removed$id & episode_id %in% inclusion_codes_with_exclusions_removed$episode_id,
    has_inclusion_but_also_exclusion_codes = id %in% intersect_between_inclusions_and_exclusions$id & episode_id %in% intersect_between_inclusions_and_exclusions$episode_id,
    has_inclusion_but_not_additional_requirements = id %in% admissions_with_inclusion_but_additional_requirements_not_satisfied$id & episode_id %in% admissions_with_inclusion_but_additional_requirements_not_satisfied$episode_id
  )

write_csv(admissions_tagged_with_inclusions, path = "../Collaborations/translated_to_icd10cm/sample_set_of_simulated_3_yrs_of_encounters_tagged_with_schnitzer2011_inclusions.csv")

## also write reduced file that's smaller
write_csv(
  admissions_tagged_with_inclusions %>% select(id, episode_id, cm, cy, meets_all_criteria, has_inclusion_but_also_exclusion_codes, has_inclusion_but_not_additional_requirements), 
  path = "../Collaborations/translated_to_icd10cm/reduced_sample_set_of_simulated_3_yrs_of_encounters_tagged_with_schnitzer2011_inclusions.csv")

```

